---
title: "CTD data"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: FALSE
    code_folding: show
    number_sections: TRUE
---

# SET-UP
CTD data for 2020 were saved every 6 seconds.

```{r markdown, echo=F}
library(knitr)
opts_chunk$set(echo = T, collapse = T, fig.align = "center", fig.width = 9, fig.height = 6)
options(width = 140)
```    

```{r settings, message = F}
source('0.0_settings.R')
```

# STATION INFO
## READ
If future years have the same file format they can simply be added in the data folder.
```{r icht_read, message = F}
# read
files <- dir(dir.icht, pattern = ".xlsx",all.files=FALSE,full.names = TRUE)
load <- lapply(files, read_excel,sheet=2)   
info <- do.call("rbind",load)

# exclude the cancelled ones
info <- info[!is.na(info$Heure),]

# format time stuff
info$day <- format(as.POSIXct(info$Date,format='%Y-%m-%d',tz='UTC'),format='%d-%m-%y') 
info$time <- format(info$Heure, "%H:%M") 
info$duration <- format(info$Duree, "%H:%M")
info$start <- as.POSIXct(paste(info$day,info$time,sep=" "),format='%d-%m-%y %H:%M',tz='UTC') #idem to ctd data
info$end <- info$start+info$`Durée totale (sec)`
```

# CTD DATA
## READ
If future years have the same file format they can simply be added in the data folder.
```{r ctd_read, message = F}
files <- dir(dir.ctd, pattern = ".xlsx",all.files=FALSE,full.names = TRUE)
load <- lapply(files, read_excel, sheet=3,skip=3,col_names=FALSE)   
col <- lapply(files, read_excel, sheet=3,skip=1,n_max=1,col_names=FALSE)  
dummy <- lapply(1:length(load),function(x)(colnames(load[[x]])<<-col[[x]][1,]))
ctd <- do.call("rbind",load)

# reformat date and time
ctd$UTC <- as.POSIXct(ctd$UTC,format='%Y-%m-%d %H:%M:%OS',tz='UTC')
ctd$day <- format(ctd$UTC,format='%d-%m-%y')
ctd$time <- format(ctd$UTC,format='%H:%M:%S')
```

## CUT
There are two methods to determine the start and stop of a CTD cast.

1. Use the start and end time as indicated onboard (when the bongo hits the water). 
  - For this mission this might not have been very precise because we didn't actually see the bongo go into the water.
  - The start time also misses the seconds (just hour and minute)

2. By checking the CTD depth and when it starts to plunge. 
  - An arbitrary rule needs to be defined to detect a plunge. Here I decided that if within ~15 seconds before or after an exact moment in time, there is an observation of a depth >1m, we can assume that we are at a new station and this is a new cast. 
  
So far I only used the second method, assuming that it should be the most precise (error of 15 seconds rather than maybe one minute).

```{r ctd_treat2, message = F}
# plunge detection (looking over 5 lines of data or 30 seconds whether in that time frame there is depth>1m)
ctd$station <- rollapply(ctd$Depth,5,function(x)any(x>1),fill=FALSE,align='center') 

# only keep station data
ctd.station <- ctd[ctd$station,] 

# give a unique id for each station (continuous series)
id <- 1
ctd.station$id <- rollapply(ctd.station$UTC,2,function(x){
  d <- difftime(x[2],x[1],units="secs",tz="UTC") # difference in time between two consecutive rows
  if(as.numeric(d)!=6)id <<- id+1 # if there are more than 6 seconds between two rows, indicate this is a new cast
  return(id)
},align='right',fill=1)
```

### PLOT {.tabset}
#### CTD profiles 1-20
Profiles allow to validate, amongst other things, that that the start and end of the cast are reasonably well identified.
```{r ctd_check_cut1,width=20,height=20}
ggplot(ctd.station[ctd.station$id %in% 1:20,],aes(x=UTC,y=Depth))+
  geom_line()+
  facet_wrap(~id,scale='free')
```

#### CTD profiles 21-40
```{r ctd_check_cut2,width=20,height=20}
ggplot(ctd.station[ctd.station$id %in% 21:40,],aes(x=UTC,y=Depth))+
  geom_line()+
  facet_wrap(~id,scale='free')
```

#### CTD profiles 41-60
```{r ctd_check_cut3,width=20,height=20}
ggplot(ctd.station[ctd.station$id %in% 41:60,],aes(x=UTC,y=Depth))+
  geom_line()+
  facet_wrap(~id,scale='free')
```

#### CTD profiles 61-80
```{r ctd_check_cut4,width=20,height=20}
ggplot(ctd.station[ctd.station$id %in% 61:80,],aes(x=UTC,y=Depth))+
  geom_line()+
  facet_wrap(~id,scale='free')
```

# CTD + STATION INFO
## Combine
```{r ctd_merge_stationinfo, message = F}
# only get essential station info
stations <- info[info$`Bab./Trib.`=='T',c('Cons','Station','start','Prof. station (m)')] 

# get the start and end of each cast
casts <- ddply(ctd.station,c('id'),summarise,time.start=min(UTC),time.end=max(UTC))  

# the cast of the station if the one where the noted down start if within or around the ctd period.
# This does a gradual search of the cast nearest in time.
stations$id <- sapply(stations$start,function(x){
      error <- 0
      y <- NULL
      while(error<(60*120) & length(y)!=1){
        y <- which((casts$time.start-error)<x & (casts$time.end+error)>x) 
        error <- error+60*5
      }
      return(y)}) 

stations <- merge(stations,casts,all.x=TRUE)
stations$bottom <- as.numeric(stations$'Prof. station (m)')

# full version
ctd.full <- merge(ctd.station,stations[,c('id','Cons','Station','bottom')])
ctd.full <- ctd.full[order(ctd.full$UTC),]

# data.frame for Melanie
ctd.summary <- ddply(ctd.full,c('Station','Cons','day','bottom'),summarise,
                 time.start=min(UTC),
                 time.end=max(UTC),
                 duration=difftime(max(UTC),min(UTC)),
                 sst10=mean(TE90[Depth<10]),
                 depth.ctd=max(Depth))
write.csv(ctd.summary,file='csv/ctd_summary.csv')
```


## PLOTS

### Profiles {.tabset}
#### Consecutives 1-22
```{r ctd_profile1}
ggplot(ctd.full[ctd.full$Cons %in% 1:22,], aes(x= UTC,y=Depth)) + 
      geom_line() +
      facet_wrap(~Cons,scale='free')+
      scale_y_reverse()
```

#### Consecutives 23-44
```{r ctd_profile2}
ggplot(ctd.full[ctd.full$Cons %in% 23:44,], aes(x= UTC,y=Depth)) + 
      geom_line() +
      facet_wrap(~Cons,scale='free')+
      scale_y_reverse()
```

#### Consecutives 45-62
```{r ctd_profile3}
ggplot(ctd.full[ctd.full$Cons %in% 45:62,], aes(x= UTC,y=Depth)) + 
      geom_line() +
      facet_wrap(~Cons,scale='free')+
      scale_y_reverse()
```

### Temperatures {.tabset}
#### all
```{r sst_all}
ggplot(ctd.full, aes(x=TE90,y=Depth,color=Cons,group=Cons)) + 
      geom_line() +
      scale_y_reverse()
```

### Bongo Depth checks {.tabset}
#### Depths
How well did we sample surface-50m?
```{r depth_check}
ggplot(ctd.summary, aes(x=bottom,y=depth.ctd)) + 
  geom_point()+
  geom_text(aes(label=Station),size=2,hjust=-0.1,vjust=-0.1)+
  labs(x='Depth of station',y='Max bongo Depth (CTD)')+
  geom_abline(slope = 1,intercept=0)+
  geom_hline(yintercept=50)
```

#### Depth margin
How far did we stay from the bottom? 
```{r depth_margin,width=6,height=5}
ctd.summary$margin <- with(ctd.summary,pmin(bottom,50)-depth.ctd)
ggplot(ctd.summary, aes(x=margin)) + 
  geom_histogram()+
  labs(x='Distance from bottom or 50m (m)')
```


